### 2019-07-18 methodology/novel-contributions.tex
hamming, jaccard distances
MPE comparison algorithm

\subsubsection{MPE algorithms comparison}

>>>>

\subsubsection{MPE algorithms comparison}
In order to compare the quality of the solution found using the simple Pseudo-MPE heuristic, I set up an experiment to compare it to the exact MPE solutions calculated with DAOOPT (see \ref{subsec:libraries}).
The user selects the number of iterations over which to average the results and at each iteration a Pseudo-MPE solution is generated using Alg. \ref{alg:pseudo-mpe-random-evidence}.
The same initial evidence is fed to DAOOPT using the \texttt{daoopt\_solver} interfacing procedure I wrote.

Returning to the example used while presenting DAOOPT in \ref{subsec:libraries}, given the \texttt{.uai} representing the BN and the \texttt{.uai.evid} random evidence:
\begin{verbatim}
 2
  4 1
  3 2
\end{verbatim}
DAOOPT would give the following output:
\begin{verbatim}
	--- Starting search ---
[0] u 3 4 -1.3581 5 2 1 2 2 1
[0] Cache statistics: . . . .

--------- Search done ---------
Problem name:  pomegranate
OR nodes:      3
AND nodes:     4
OR processed:  3
AND processed: 8
Leaf nodes:    2
Pruned nodes:  4
Deadend nodes: 1
Time elapsed:  0 seconds
Preprocessing: 0 seconds
-------------------------------
-1.3581 (0.0438433)

p 2 1 2
l 2 1 6
s -1.3581 5 2 1 2 2 1
\end{verbatim}
The end of the final line is the one of interest as it is the assignment of values to the variables that solves the MPE problem.
The \texttt{5 2 1 2 2 1} string is to be interpreted as meaning:
\begin{itemize}
  \item there are 5 variables in the solution
  \item the variable indexed by 0 (in the ordering given in the preable of the \texttt{.uai}) is assigned its second value (the ordering is inferred by the CPTs defined in the \texttt{.uai}) in the MPE solution
  \item variable 1 is assigned its first value
  \item variable 2 is assigned its second value
  \item variable 3 is assigned its second value
  \item variable 4 is assigned its first value
\end{itemize}
Variables 3 and 4 are constrained to assume the value specified in the input \texttt{.uai.evid}; in this case 1 and 2, respectively.

At each iteration corresponds a random evidence; the solutions given by the Pseudo-MPE and by DAOOPT are scored against each other using Hamming (Subsec. \ref{subsec:hamming-distance}) and Jaccard (Subsec. \ref{subsec:jaccard-distance}) distances on the vectors representing the returned assignments.
The sequence of distances is then averaged over the iterations to return the two averaged distances between the Pseudo-MPE and DAOOPT solutions to the MPE problem.

The implementation of the algorithm had to deal with the fact that DAOOPT failed on certain configurations of input evidence.
The reason for this is unknown so my workaround was to discard any iterations where DAOOPT had failed and retry them with another random set of evidences.
The pseudocode for the algorithm is shown in Alg. \ref{alg:mpe-comparison}.

\begin{algorithm}[htp!]
	\caption{MPE comparison algorithm}
	\label{alg:mpe-comparison}
	\begin{algorithmic}[1]
		\State user defined number of $iterations$
		\State initialise empty $results$ list
		\While{$i < iterations$}
			\State generate random number $k \in [ 1, |V| ]$
			\State $S = $ choose $k$ variables from $V$
			\For{$s \in S$}
				\State choose random $v$ in the possible values of $e$ 
				\State $evidence = evidence \cup (s,v)$
			\EndFor
			\State $pseudo\_mpe = $ \texttt{generate\_pseudo\_mpe($evidence$)}
			\State $model\_uai=$\texttt{export\_model\_to\_uai()}
			\State $evidence\_uai=$\texttt{export\_evidence\_to\_uai()}
			\State $daoopt\_mpe = $ \texttt{daoopt\_solver($model\_uai$, $evidence\_uai$)}
			\If{$daoopt\_mpe$ == -1} \Comment{DAOOPT has failed to calculate a solution}
				\State cancel $pseudo\_mpe$
				\State continue
			\EndIf
			\State compute $d_H(pseudo\_mpe,daoopt\_mpe)$, $d_J(pseudo\_mpe,daoopt\_mpe)$ and store in $iter\_result$
			\State append $iter\_result$ to $results$
			\State $i+=1$
		\EndWhile
		\State compute average in $results$ for $d_H$ and $d_J$
	\end{algorithmic}
\end{algorithm}
